---
title: "Compare EC50 models"
author: "Kaique S Alves"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Compare EC50 models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

Use `ec50_multimodel()` when you want to fit more than one candidate `drc` model to the same isolates and strata. This is useful during model selection, when you need EC estimates and model-selection statistics in the same output.

## Packages

```{r message=FALSE, warning=FALSE}
library(ec50estimator)
library(drc)
library(ggplot2)
```

## Data

The examples use a small subset of `multi_isolate` so the vignette runs quickly while still showing the grouped workflow.

```{r}
data(multi_isolate)

example_data <- subset(
  multi_isolate,
  isolate %in% 1:5 & fungicida == "Fungicide A"
)

head(example_data)
```

Inspect the raw dose-response observations before comparing models. Because each isolate can have repeated observations at the same dose, this plot uses jittered points instead of connecting observations with lines.

```{r fig.alt="Jittered dose-response scatter plot for five isolates exposed to Fungicide A, faceted by field system."}
ggplot(example_data, aes(dose, growth, color = factor(isolate))) +
  geom_jitter(width = 0.08, height = 0, alpha = 0.85, size = 2) +
  facet_wrap(~field) +
  scale_x_log10() +
  labs(x = "Dose", y = "Growth", color = "Isolate") +
  theme_light()
```

## Compare Models

Pass a list of candidate model functions to `fct`. The function fits each model for each isolate and stratum, then appends model-selection statistics returned by `drc::mselect()`.

```{r}
df_models <- ec50_multimodel(
  growth ~ dose,
  data = example_data,
  isolate_col = "isolate",
  strata_col = "field",
  fct = list(drc::LL.3(), drc::LL.4(), drc::W2.3()),
  interval = "delta"
)

head(df_models)
```

The `model` column identifies the candidate model. The `IC` column is the information criterion reported by `drc::mselect()` for the fitted model.

## Pick Candidate Models

You can rank candidate models within each isolate and stratum using base R. Lower `IC` values indicate stronger relative support among the compared models.

```{r}
ranked_models <- df_models[order(df_models$field, df_models$ID, df_models$IC), ]
best_models <- ranked_models[!duplicated(ranked_models[c("field", "ID")]), ]

best_models[, c("ID", "field", "model", "Estimate", "Std..Error", "IC")]
```

## Visualize Model Comparison

Plot the information criterion across isolates to see whether one model is consistently preferred.

```{r fig.alt="Point plot of information criterion values by isolate for each candidate model and field system."}
plot_models <- df_models
plot_models$ID <- as.numeric(plot_models$ID)

ggplot(plot_models, aes(ID, IC, color = model)) +
  geom_point(position = position_dodge(width = 0.35), size = 2) +
  facet_wrap(~field) +
  scale_x_continuous(breaks = sort(unique(plot_models$ID))) +
  labs(x = "Isolate", y = "Information criterion") +
  theme_light()
```

After choosing an appropriate model, use `estimate_EC50()` for the final estimation workflow if a single model should be applied across the full dataset.
