---
title: "Estimate EC50 with one model"
author: "Kaique S Alves"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Estimate EC50 with one model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

Use `estimate_EC50()` when you already know which dose-response model should be fitted to each isolate. The function runs the same `drc` model across isolates and optional strata, then returns a single data frame with the EC estimates.

## Packages

```{r message=FALSE, warning=FALSE}
library(ec50estimator)
library(drc)
library(ggplot2)
library(cowplot)
library(ggridges)
```

## Data

The package includes `multi_isolate`, a simulated dataset of mycelial growth under increasing fungicide doses for 50 fungal isolates, two field systems, and two fungicides.

```{r}
data(multi_isolate)
head(multi_isolate)
```

Inspect the raw dose-response data before fitting models.

```{r fig.alt="Jitter plot of mycelial growth by fungicide dose, faceted by field system and fungicide."}
ggplot(multi_isolate, aes(dose, growth, color = growth)) +
  geom_jitter(width = 0.1) +
  facet_grid(field ~ fungicida) +
  scale_x_log10() +
  scale_color_viridis_c(option = "B", direction = 1) +
  theme_light()
```

## Estimate EC50

`estimate_EC50()` needs the same model ingredients you would pass to `drc::drm()`: a formula, a data frame, and a model function such as `drc::LL.3()`. Use `isolate_col` to identify isolates and `strata_col` for grouping columns such as field, fungicide, year, or location.

```{r}
df_ec50 <- estimate_EC50(
  growth ~ dose,
  data = multi_isolate,
  isolate_col = "isolate",
  strata_col = c("field", "fungicida"),
  interval = "delta",
  fct = drc::LL.3()
)

head(df_ec50)
```

The result keeps the isolate identifier in `ID`, then returns the strata columns and the effective-dose estimate columns from `drc::ED()`.

## Visualize Results

Plot EC50 estimates and confidence intervals for all isolates.

```{r fig.height=10, fig.width=6, fig.alt="Point and interval plot of EC50 estimates for isolates, grouped by field system and fungicide."}
plot_ec50 <- df_ec50
plot_ec50$ID <- as.numeric(plot_ec50$ID)

ggplot(plot_ec50, aes(ID, Estimate, color = field)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Lower, ymax = Upper, color = field), width = 0) +
  facet_wrap(~fungicida, scales = "free_x", ncol = 2) +
  scale_y_log10() +
  scale_x_continuous(breaks = 1:50) +
  scale_color_manual(values = c("darkred", "steelblue")) +
  labs(x = "Isolates", y = "EC50") +
  theme_minimal_vgrid(font_size = 10) +
  coord_flip() +
  theme(
    axis.text.x = element_text(size = 10),
    legend.position = "bottom"
  )
```

Density plots can help compare EC50 distributions between strata.

```{r fig.alt="Density ridge plot comparing EC50 distributions between field systems for each fungicide."}
ggplot(df_ec50, aes(Estimate, field, fill = after_stat(x))) +
  geom_density_ridges_gradient(alpha = 0.3) +
  scale_x_log10() +
  scale_fill_viridis_c(option = "C") +
  facet_wrap(~fungicida, nrow = 2) +
  theme_minimal_hgrid() +
  labs(x = "EC50", y = "Field") +
  theme(legend.position = "none")
```
